---
title: "runneR Data Validation"
author: "Samuel Ivanecky"
date: '2022-06-13'
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=18, fig.height=10) 

# Code to generate line item performances from TFRRS.
library(tidymodels)
library(httr)
library(jsonlite)
library(RPostgreSQL)
library(DBI)
library(reshape2)
library(stringr)
library(stringi)
library(yaml)
library(rvest)
library(kit)
library(ggthemes)
library(ggridges)
library(ggrepel)
library(ggpubr)
library(zoo)
library(gt)
library(gtExtras)
library(data.table)
library(tidytable)
library(dplyr)

# Source file for functions
source("/Users/samivanecky/git/runneR/scrapeR/Scraping_Fxns.R")
source("/Users/samivanecky/git/runneR/scrapeR/meetScrapingFxns.R")
source("/Users/samivanecky/git/runneR/scrapeR/altConversinFxns.R")

# Read in conversion data for track size
trckSz <- read.csv("/Users/samivanecky/git/runneR/ncaa_size_conversion.csv")
trckSz <- trckSz %>%
  mutate(
    gender = case_when(
      gender == "men" ~ "M",
      T ~ "F"
    )
  )

# Read connection data from yaml
pg.yml <- read_yaml("/Users/samivanecky/git/postgres.yaml")

# Connect to database
pg <- dbConnect(
  RPostgres::Postgres(),
  host = pg.yml$host,
  user = pg.yml$user,
  db = pg.yml$database,
  port = pg.yml$port
)

reformatTimes <- function(mark) {
  sec <- str_pad((mark %% 60), width = 2, side = "left", pad = "0")
  min <- floor(mark / 60)
  time <- paste0(min, ":", sec)
  return(time)
}
```

# Query Data
```{r}
# Query
results <- dbGetQuery(pg, "select distinct * from meet_results_raw")

cal <- genCal()
```

# Clean Data
## Remove Rows with NA
```{r}
# Subset NA data
na_rows <- results %>%
  filter(
    is.na(NAME) | is.na(TIME) | is.na(RACE_NAME)
  )

# Remove NA rows
res_sub <- results %>%
  filter(
    !(is.na(NAME) | is.na(TIME) | is.na(RACE_NAME))
  )
```

## Clean Up Fields
```{r}
# Convert to data.table
res_sub <- data.table::as.data.table(res_sub)

res_sub <- res_sub %>%
  mutate.(
    NAME = str_squish(NAME),
    RACE_NAME = str_squish(RACE_NAME)
  ) %>%
  mutate.(
    # Prelim
    is_prelim = case_when(
      grepl("prelim|heats", RACE_NAME, ignore.case = TRUE) ~ 1,
      grepl("prelim", MEET_NAME, ignore.case = TRUE) ~ 1,
      T ~ 0
    ),
    # Race distance
    event = case_when(
      grepl("10000 meter|10k|10,000|10000", RACE_NAME, ignore.case = TRUE) & !grepl("xc|cross country|cc", RACE_NAME, ignore.case = TRUE) ~ "10k",
      grepl("5000 meter|10k|5,000|5000", RACE_NAME, ignore.case = TRUE) & !grepl("xc|cross country|cc", RACE_NAME, ignore.case = TRUE) ~ "5k",
      grepl("10000 meter|10k|10,000|10000", RACE_NAME, ignore.case = TRUE) & (grepl("xc|cross country|cc", RACE_NAME, ignore.case = TRUE) | grepl("xc|cross country|cc", MEET_NAME, ignore.case = TRUE)) ~ "10k XC",
      grepl("5000 meter|10k|5,000|5000", RACE_NAME, ignore.case = TRUE) & (grepl("xc|cross country|cc", RACE_NAME, ignore.case = TRUE) | grepl("xc|cross country|cc", MEET_NAME, ignore.case = TRUE)) ~ "5k XC",
      grepl("6000m|6000|6,000|6k", RACE_NAME, ignore.case = TRUE) & (grepl("xc|cross country|cc", RACE_NAME, ignore.case = TRUE) | grepl("xc|cross country|cc", MEET_NAME, ignore.case = TRUE)) ~ "6k XC",
      grepl("100 meter|100m|100 m", RACE_NAME, ignore.case = TRUE) & !grepl("hurdle|mH", RACE_NAME, ignore.case = TRUE) ~ "100m",
      grepl("200 meter|200m|200 m", RACE_NAME, ignore.case = TRUE) ~ "200m",
      grepl("400 meter|400m|400 m", RACE_NAME, ignore.case = TRUE) & !grepl("hurdle|mH", RACE_NAME, ignore.case = TRUE) ~ "400m",
      grepl("800 meter|800m|800 m", RACE_NAME, ignore.case = TRUE) ~ "800m",
      grepl("1500 meter|1500m|1500 m|1,500", RACE_NAME, ignore.case = TRUE) ~ "1500m",
      grepl("mile", RACE_NAME, ignore.case = TRUE) ~ "Mile",
      grepl("3000|3k|3000m|3000 m|3,000", RACE_NAME, ignore.case = TRUE) & !grepl("3000S|steeple", RACE_NAME, ignore.case = TRUE) ~ "3000m",
      (grepl("3000|3000m|3000 m", RACE_NAME, ignore.case = TRUE) & grepl("steeple", RACE_NAME, ignore.case = TRUE)) | grepl("3000S", RACE_NAME, ignore.case = TRUE) ~ "3000S",
      grepl("100 meter|100m|100 hurdles|100 m hurdle|100m hurdle", RACE_NAME, ignore.case = TRUE) & grepl("hurdle|mH|100H", RACE_NAME, ignore.case = TRUE) ~ "100mH",
      grepl("110 meter|110m|110 hurdles|110 m hurdle|110m hurdle", RACE_NAME, ignore.case = TRUE) & grepl("hurdle|mH|110H", RACE_NAME, ignore.case = TRUE) ~ "110mH",
      grepl("400 meter|400mH|400 hurdles|400 m hurdle|400m hurdle", RACE_NAME, ignore.case = TRUE) & grepl("hurdle|mH|400H|400mH", RACE_NAME, ignore.case = TRUE) ~ "400mH",
      grepl("60 meter|60m|60 m", RACE_NAME, ignore.case = TRUE) & !grepl("hurdle|60mH", RACE_NAME, ignore.case = TRUE) ~ "60m",
      grepl("60 meter|60m| 60 hurdles|60 m hurdle", RACE_NAME, ignore.case = TRUE) & grepl("hurdle|60mH", RACE_NAME, ignore.case = TRUE) ~ "60mH",
      grepl("55 meter|55m| 55 hurdles|55 m hurdle", RACE_NAME, ignore.case = TRUE) & grepl("hurdle|55mH", RACE_NAME, ignore.case = TRUE) ~ "60mH",
      grepl("55 meter|55m|55 m", RACE_NAME, ignore.case = TRUE) & !grepl("hurdle", RACE_NAME, ignore.case = TRUE) ~ "55m",
      grepl("600 meter|600m|600 m", RACE_NAME, ignore.case = TRUE) ~ "600m",
      grepl("1000 meter|1000m|1k|1000 m", RACE_NAME, ignore.case = TRUE) ~ "1000m",
      grepl("300 meter|300m|300 m", RACE_NAME, ignore.case = TRUE) ~ "300m",
      grepl("500 meter|500m|500 m", RACE_NAME, ignore.case = TRUE) ~ "500m",
      grepl("2000 steeplechase", RACE_NAME, ignore.case = TRUE) | (grepl("2000", RACE_NAME) & grepl("steeple", RACE_NAME, ignore.case = TRUE)) ~ "2000S",
      grepl("1500 steeplechase", RACE_NAME, ignore.case = TRUE) | (grepl("1500", RACE_NAME) & grepl("steeple", RACE_NAME, ignore.case = TRUE)) ~ "1500S",
      T ~ "OTHER"
    )
  )
```

## Break Out Other Events
```{r}
# Filter out OTHER events
other_evnts <- res_sub %>%
  filter.(
    event == "OTHER"
  )

res_sub <- res_sub %>%
  filter.(
    event != "OTHER"
  )

# Define gender
res_sub <- res_sub %>%
  mutate.(
    gender = case_when(
      grepl("men|male|boy", RACE_NAME, ignore.case = TRUE) & !grepl("women", RACE_NAME, ignore.case = TRUE) ~ "M",
      grepl("women|woman|girl|female", RACE_NAME, ignore.case = TRUE) ~ "F",
      T ~ NA
    )
  )
```

## Create Runner Lookup Dictionary
```{r}
# Lookup dictionary
rnrs <- res_sub %>%
  select.(
    NAME, TEAM, YEAR, gender
  ) %>%
  funique() %>%
  mutate.(
    class = case_when(
      grepl("fr|freshman", YEAR, ignore.case = TRUE) ~ "FR",
      grepl("so|sophomore", YEAR, ignore.case = TRUE) ~ "SO",
      grepl("jr|junior", YEAR, ignore.case = TRUE) ~ "JR",
      grepl("sr|senior", YEAR, ignore.case = TRUE) ~ "SR",
      grepl("un|unattached", YEAR, ignore.case = TRUE) | grepl("unattached", TEAM, ignore.case = TRUE) ~ "Unattached",
      T ~ "OTHER"
    )
  )

# Split out non-year runners
no_year <- rnrs %>%
  filter.(
    class == "OTHER"
  ) 

rnrs <- rnrs %>%
  filter.(
    class != "OTHER"
  ) %>%
  summarise.(
    gender = max(gender, na.rm = T),
    .by = c(NAME, TEAM, YEAR, class)
  ) %>%
  funique()
```

## Correct Class in Data
```{r}
# Drop current year and join in data
res_ <- res_sub %>%
  select.(-c(gender)) %>%
  left_join.(rnrs, by = c("NAME", "TEAM", "YEAR")) %>%
  select(-c(YEAR))
```

## Get Meet Dates
```{r}
res_ <- res_ %>%
  mutate.(
    month = case_when(
      grepl("January", MEET_DATE, ignore.case = TRUE) ~ "1",
      grepl("February", MEET_DATE, ignore.case = TRUE) ~ "2",
      grepl("March", MEET_DATE, ignore.case = TRUE) ~ "3",
      grepl("April", MEET_DATE, ignore.case = TRUE) ~ "4",
      grepl("May", MEET_DATE, ignore.case = TRUE) ~ "5",
      grepl("June", MEET_DATE, ignore.case = TRUE) ~ "6",
      grepl("July", MEET_DATE, ignore.case = TRUE) ~ "7",
      grepl("August", MEET_DATE, ignore.case = TRUE) ~ "8",
      grepl("September", MEET_DATE, ignore.case = TRUE) ~ "9",
      grepl("October", MEET_DATE, ignore.case = TRUE) ~ "10",
      grepl("Novemeber", MEET_DATE, ignore.case = TRUE) ~ "11",
      grepl("December", MEET_DATE, ignore.case = TRUE) ~ "12",
      T ~ 'OTHER'
    ),
    year = trimws(substr(MEET_DATE, nchar(MEET_DATE) - 4, nchar(MEET_DATE))),
    day = trimws(stri_extract_first_regex(MEET_DATE, "[0-9]+"))
  ) %>%
  mutate.(
    race_date = lubridate::ymd(paste0(year, "-", month, "-", day))
  )
```

## Get Numeric Times
```{r}
# Convert times
res_$numeric_time = sapply(res_$TIME, handleTimes)
res_ <- res_ %>%
  mutate.(
    numeric_time = as.numeric(numeric_time)
  ) %>%
  ungroup()

# Convert back to dataframe
# res_ <- as.data.frame(res_)
```

## Filter Out Gender-less Data
```{r}
res_ <- res_ %>%
  filter(!is.na(gender)) %>%
  mutate(
    MEET_DATE = trimws(MEET_DATE),
    MEET_LOCATION = trimws(MEET_LOCATION),
    MEET_NAME = trimws(MEET_NAME)
  )
```

## Load & Handle Meet Lookup Data
```{r}
# Query meet dates
meet_lookup <- dbGetQuery(pg, "select distinct * from meet_lookup_info") %>%
  select(-c(race_date)) %>%
  mutate(
    join_key = paste0(gsub("[[:space:]]", "", name), '-', gsub("[[:space:]]", "", meet_date))
  )

# Join to results data
res_ <- res_ %>%
  mutate(
    join_key = paste0(gsub("[[:space:]]", "", MEET_NAME), '-', gsub("[[:space:]]", "", MEET_DATE))
  ) %>%
  left_join.(meet_lookup, by = c("join_key"))
```

## Perform Time Conversions
```{r}
# Drop unused columns
res_ <- res_ %>%
  select(-c(load_d, join_key, dets, name, meet_date))

# Clean up facility fields
res_$elevation = ifelse(is.na(res_$elevation), 0, res_$elevation)
res_$track_size = ifelse(is.na(res_$track_size), "400m", res_$track_size)
res_$banked_or_flat = ifelse(is.na(res_$banked_or_flat), "flat", res_$banked_or_flat)
res_$elevation = ifelse(res_$elevation == "", 0, res_$elevation)
res_$track_size = ifelse(res_$track_size == "", "400m", res_$track_size)
res_$banked_or_flat = ifelse(res_$banked_or_flat == "", "flat", res_$banked_or_flat)
```

### Handle "Bad" Meet Data from Scraping
```{r}
res_ <- res_ %>%
  mutate.(
    banked_or_flat = case_when(
      !(month %in% c("11", "12", "1", "2", "3")) ~ 'flat',
      T ~ banked_or_flat
    ),
    track_size = case_when(
      !(month %in% c("11", "12", "1", "2", "3")) ~ '400m',
      T ~ track_size
    )
  )
```

```{r}
res_ <- res_ %>%
  mutate.(
    track_size = as.numeric(gsub("m", "", track_size))
  ) %>%
  mutate.(
    track_size_conversion = case_when(
      track_size < 200 & event %in% c("800m", "Mile", "1500m", "3000m", "5000m", "400m", "DMR", "1000m", "600m", "500m") ~ "undersized",
      track_size == 200 & banked_or_flat == "flat" & event %in% c("800m", "Mile", "1500m", "3000m", "5000m", "400m", "DMR", "1000m", "600m", "500m", "200m", "300m") ~ "flat",
      T ~ "none"
    )
  ) %>%
  mutate.(
    elevation = ifelse(is.na(as.numeric(elevation)), 0, as.numeric(elevation))
  ) %>%
  filter(!is.na(elevation))
```

```{r}
# Generate altitude converted mark
res_$alt_conv_mark <- mapply(getConv, alt = res_$elevation, event = res_$event, mark = res_$numeric_time)

# Generate track size converted mark
# Join track size fields for conversion
res_ <- res_ %>%
  left_join.(
    trckSz, by = c("event", "gender", "track_size_conversion" = "type")
  )

# Create final converted mark (alt + track)
res_$conversion = ifelse(is.na(res_$conversion), 1, res_$conversion)

# Create converted time
res_ <- res_ %>%
  mutate.(
    converted_time = conversion * alt_conv_mark
  )

# Round data
res_ <- res_ %>%
  mutate.(
    converted_time = round(converted_time, 2)
  )
```

## Drop Unused Data
```{r}
res_ <- res_ %>%
  select(-c(facility, conversion, load_date, alt_conv_mark, track_size_conversion))
```

## Create New Fields
```{r}
# Determine what season (indoor vs outdoor)
res_ <- res_ %>%
  mutate(
    season_cat = case_when(
     track_size != "400" ~ "indoor",
      T ~ "outdoor"
    )
  )
```

## Upload to Database
```{r}
# Convert back to data frame
res_ <- as.data.frame(res_)

# Upload
dbRemoveTable(pg, "meet_results_dets")
dbCreateTable(pg, "meet_results_dets", res_)
dbWriteTable(pg, "meet_results_dets", res_, append = TRUE)
```

###############################
# Data Validation
## Overall Counts by Race / Gender / Year
```{r}
# Get counts of event by gender by year
race_cnts <- res_ %>%
  summarise.(
    n_races = n(),
    .by = c(year, gender, event)
  )
```

# Data Analysis
## Check Top Marks by Year
### Steeplechase
```{r}
res_ %>%
  filter(gender == "F") %>%
  filter(event == "3000S") %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Women's Top 100 Steeplechase Times by Season",
    subtitle = "Showing top 100 marks based on competition season. Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) +
  scale_x_continuous(name = "Time",
                     breaks = c(550, 560, 570, 580, 590, 600, 610, 620, 630, 640, 650, 660),
                     labels = c("9:10", "", "9:30", "", "9:50", "", "10:10", "", "10:30", "", "10:50", ""))

res_ %>%
  filter(gender == "M") %>%
  filter(event == "3000S") %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Men's Top 100 Steeplechase Times by Season",
    subtitle = "Showing top 100 marks based on competition season. Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) +
  scale_x_continuous(name = "Time",
                     breaks = c(490, 500, 510, 520, 530, 540, 550, 560, 570, 580, 590, 600),
                     labels = c("8:10", "", "8:30", "", "8:50", "", "9:10", "", "9:30", "", "9:50", ""))
```

### 5k
```{r}
res_ %>%
  dplyr::filter(
      year != "2009" &
      gender == "F" &
      event == "5k" &
      TEAM != "Unattached"
  ) %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Women's Top 100 5k Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) +
  scale_x_continuous(name = "Time",
                     breaks = c(900, 920, 940, 960, 980, 1000, 1020),
                     labels = c("15:00", "", "15:40", "", "16:20", "", "17:00"))

res_ %>%
  filter(gender == "M") %>%
  filter(event == "5k") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Men's Top 100 5k Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) +
  scale_x_continuous(name = "Time",
                     breaks = c(780, 790, 800, 810, 820, 830, 840, 850, 860),
                     labels = c("13:00", "", "13:20", "", "13:40", "", "14:00", "", "14:20"))
```

#### Median Times
```{r}
# Men's data
med_times_m <- res_ %>%
  filter(gender == "M") %>%
  filter(event == "5k") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  slice_min(order_by = numeric_time, n = 100) %>%
  summarise(
    median_time = median(numeric_time, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    time_shift = median_time - lag(median_time),
    rolling_avg = cummean(median_time)
  ) %>%
  mutate(
    time_shift_v_cavg = median_time - lag(rolling_avg)
  ) %>%
  mutate(
    gender = "Men"
  )

# Women's data
med_times_f <- res_ %>%
  filter(gender == "F") %>%
  filter(event == "5k") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  slice_min(order_by = numeric_time, n = 100) %>%
  summarise(
    median_time = median(numeric_time, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    time_shift = median_time - lag(median_time),
    rolling_avg = cummean(median_time)
  ) %>%
  mutate(
    time_shift_v_cavg = median_time - lag(rolling_avg)
  ) %>%
  mutate(
    gender = "Women"
  )

# Combine data
med_times <- rbind(med_times_f, med_times_m)

# Plot
med_times %>%
  mutate(
    time_shift_v_cavg = -(time_shift_v_cavg)
  ) %>%
  ggplot(aes(year, time_shift_v_cavg, color = gender, group = gender)) +
  geom_point(size = 8) +
  geom_line(size = 3) +
  ggplot2::labs(
    x = "Competition Year",
    y = "Time Shift vs Historic Average (sec)",
    caption = "",
    title = "Median 5k Time Shift of Top 100 Performances by NCAA Season",
    subtitle = "Using top 100 marks by season for each year. Time shift is median time vs historic average. Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  )
```

## 100m
```{r}
res_ %>%
  dplyr::filter(
      year != "2009" &
      gender == "F" &
      event == "100m" &
      TEAM != "Unattached"
  ) %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Women's Top 100 100m Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) #+
  #scale_x_continuous(name = "Time",
  #                   breaks = c(900, 920, 940, 960, 980, 1000, 1020),
  #                   labels = c("15:00", "", "15:40", "", "16:20", "", "17:00"))

res_ %>%
  filter(gender == "M") %>%
  filter(event == "100m") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  arrange(numeric_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(numeric_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Men's Top 100 100m Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  )# +
  #scale_x_continuous(name = "Time",
  #                   breaks = c(780, 790, 800, 810, 820, 830, 840, 850, 860),
  #                   labels = c("13:00", "", "13:20", "", "13:40", "", "14:00", "", "14:20"))
```

## Mile
```{r}
res_ %>%
  dplyr::filter(
      year > 2014 &
      gender == "F" &
      event == "Mile" &
      TEAM != "Unattached"
  ) %>%
  group_by(year) %>%
  arrange(converted_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(converted_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Women's Top 100 Mile Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  ) #+
  #scale_x_continuous(name = "Time",
  #                   breaks = c(900, 920, 940, 960, 980, 1000, 1020),
  #                   labels = c("15:00", "", "15:40", "", "16:20", "", "17:00"))

res_ %>%
  filter(gender == "M") %>%
  filter(event == "Mile") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2014) %>%
  group_by(year) %>%
  arrange(converted_time, .by_group = T) %>%
  mutate(
    rank = 1:n()
  ) %>%
  ungroup() %>%
  filter(rank <= 100) %>%
  ggplot(aes(converted_time, year, fill = year)) +
  geom_density_ridges(alpha = 0.6) +
  ggplot2::labs(
    x = "",
    y = "Competition Season",
    caption = "",
    title = "NCAA Men's Top 100 Mile Times by Season",
    subtitle = "Showing top 100 marks based on competition season (includes indoor & outdoor). Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  )# +
  #scale_x_continuous(name = "Time",
  #                   breaks = c(780, 790, 800, 810, 820, 830, 840, 850, 860),
  #                   labels = c("13:00", "", "13:20", "", "13:40", "", "14:00", "", "14:20"))
```

### Median Times
```{r}
# Men's data
med_times_m <- res_ %>%
  filter(gender == "M") %>%
  filter(event == "100m") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  slice_min(order_by = numeric_time, n = 100) %>%
  summarise(
    median_time = median(numeric_time, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    time_shift = median_time - lag(median_time),
    rolling_avg = cummean(median_time)
  ) %>%
  mutate(
    time_shift_v_cavg = median_time - lag(rolling_avg)
  ) %>%
  mutate(
    gender = "Men"
  )

# Women's data
med_times_f <- res_ %>%
  filter(gender == "F") %>%
  filter(event == "100m") %>%
  filter(TEAM != "Unattached") %>%
  filter(year > 2009) %>%
  group_by(year) %>%
  slice_min(order_by = numeric_time, n = 100) %>%
  summarise(
    median_time = median(numeric_time, na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(
    time_shift = median_time - lag(median_time),
    rolling_avg = cummean(median_time)
  ) %>%
  mutate(
    time_shift_v_cavg = median_time - lag(rolling_avg)
  ) %>%
  mutate(
    gender = "Women"
  )

# Combine data
med_times <- rbind(med_times_f, med_times_m)

# Plot
med_times %>%
  mutate(
    time_shift_v_cavg = -(time_shift_v_cavg)
  ) %>%
  ggplot(aes(year, time_shift_v_cavg, color = gender, group = gender)) +
  geom_point(size = 8) +
  geom_line(size = 3) +
  ggplot2::labs(
    x = "Competition Year",
    y = "Time Shift vs Historic Average (sec)",
    caption = "",
    title = "Median 100m Time Shift of Top 100 Performances by NCAA Season",
    subtitle = "Using top 100 marks by season for each year. Time shift is median time vs historic average. Please note data may occasionally be missing. Contact @ivanxcky if you notice anything."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold")
  )
```


## Plot Data
```{r}
race_cnts %>%
  filter(!is.na(gender)) %>%
  ggplot(aes(year, n_races, color = gender, group = gender)) +
  geom_line() +
  geom_point() +
  facet_wrap(event ~ ., scales = "free_y")
```

# Data Analysis
## Steeplechase Plot
```{r}
res_ %>%
  filter(gender == "F") %>%
  filter(event == "3000S") %>%
  ggplot(aes(numeric_time, fill = year)) +
  geom_histogram() +
  ggplot2::labs(
    x = "",
    y = "",
    caption = "",
    title = "Women's Steeplechase Times by Season",
    subtitle = ""
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 40, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=25, face = "bold")
  ) +
  facet_wrap(year ~ ., scales = "free_y")

```

##