---
title: "Ad-hoc Indoor Analysis"
author: "Samuel Ivanecky"
date: "12/4/2021"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=18, fig.height=10) 

# Code to generate line item performances from TFRRS.
library(tidymodels)
library(httr)
library(jsonlite)
library(RPostgreSQL)
library(DBI)
library(reshape2)
library(stringr)
library(yaml)
library(rvest)
library(kit)
library(ggthemes)
library(ggridges)
library(ggrepel)
library(ggpubr)
library(zoo)
library(dplyr)
library(gt)
library(gtExtras)
library(data.table)
library(tidytable)

# Source file for functions
source("/Users/samivanecky/git/runneR/scrapeR/Scraping_Fxns.R")
source("/Users/samivanecky/git/runneR/scrapeR/meetScrapingFxns.R")

# Read connection data from yaml
pg.yml <- read_yaml("/Users/samivanecky/git/runneR/postgres.yaml")

# Connect to database
pg <- dbConnect(
  RPostgres::Postgres(),
  host = pg.yml$host,
  user = pg.yml$user,
  db = pg.yml$database,
  port = pg.yml$port
)

reformatTimes <- function(mark) {
  sec <- str_pad((mark %% 60), width = 2, side = "left", pad = "0")
  min <- floor(mark / 60)
  time <- paste0(min, ":", sec)
  return(time)
}
```

# Data Setup

## Run Script to Load Data, Query Data

```{r echo=TRUE, warning=FALSE}
source("/Users/samivanecky/git/runneR/scrapeR/dataCleaneR.R")

# Remove extra data
rm(runLines)
rm(linesRunner)
rm(linesField)
rm(lines)
rm(fieldLines)
```

## Query

### Overall Data

```{r warning=FALSE}
# Query
rl <- dbGetQuery(pg, "select * from runner_lines_details")

cal <- genCal()

# Filter to unique
rl <- rl %>%
  funique()

# Try and remove duplicates
rl <- rl %>%
  filter(grepl("\\.|:", MARK) & MARK_TIME != -999) %>%
  filter(IS_FIELD == FALSE & converted_time != -999 & EVENT_TYPE %in% c("INDOOR", "OUTDOOR", "XC")) %>%
  group_by(RUNNER_KEY) %>%
  arrange(meet_date, .by_group = TRUE) %>%
  mutate(
    is_duplicate = case_when(
      MARK_TIME == lag(MARK_TIME) & EVENT == lag(EVENT) & (lubridate::ymd(meet_date) - lubridate::ymd(lag(meet_date)) < 2) ~ 'Y',
      T ~ 'N'
    )
  ) %>%
  ungroup() %>%
  group_by(RUNNER_KEY, EVENT, EVENT_TYPE) %>%
  arrange(MARK_TIME, .by_group = TRUE) %>%
  mutate(
    is_duplicate = case_when(
      MARK_TIME == lag(MARK_TIME) ~ 'Y',
      T ~ is_duplicate
    )
  ) %>%
  filter(is_duplicate == 'N')
```

### Aggregate Divisions

```{r warning=FALSE}
# Convert to DT
rlDT <- data.table::setDT(rl)

# Subset teams for joining division
teams <- rl %>%
  select(TEAM, DIVISION) %>%
  funique() %>%
  filter(DIVISION != "OTHER")

divNames <- rlDT %>%
  select.(NAME, GENDER, TEAM) %>%
  funique() %>%
  left_join.(teams, by = c("TEAM"))
```

## Recreate Grouped Yearly Data Using Competition Year

```{r warning=FALSE}
# Create all-time grouped data
grpYrly <- rlDT %>%
  filter.(IS_FIELD == FALSE) %>%
  summarize.(
    pr = min(converted_time, na.rm = T),
    prs_run = sum(is_pr, na.rm = T),
    avg_time = round(mean(converted_time, na.rm = T), 2),
    times_run = n_distinct(MEET_NAME),
    avg_place = round(mean(PLACE, na.rm = T), 2),
    wins = n_distinct(MEET_NAME[PLACE == 1]),
    .by = c(NAME, TEAM, GENDER, EVENT, competition_year)
  ) %>%
  mutate.(
    win_pct = round((wins / times_run) * 100, 3)
  )

# Rowwise operations
grpYrly$clean_pr = sapply(grpYrly$pr, reformatTimes)
grpYrly$clean_avg_time = sapply(grpYrly$avg_time, reformatTimes)

# Convert back to dataframe
grpYrly <- as.data.frame(grpYrly)

```

## Strength Of Competition
### Get Grouped Data
```{r warning=FALSE}
# Line item level
raceGrp <- rl %>%
  mutate(
    TIME = converted_time
  ) %>%
  group_by(MEET_NAME, meet_date, competition_year, GENDER, EVENT, PRELIM) %>%
  arrange(
    TIME, .by_group = TRUE
  ) %>%
  mutate(
    meet_evnt_rank = 1:n(),
    time_behind_winner = TIME - min(TIME, na.rm = T),
    place_behind_winner = PLACE - min(PLACE, na.rm = T)
  ) %>%
  select(-c(DIVISION)) %>%
  left_join(teams, by = c("TEAM"))
```

### Get Best Times for Time Durations
```{r warning=FALSE}
# Get best weekly times
raceBestWeekly <- rl %>%
  filter(is_duplicate == "N") %>%
  filter(IS_FIELD == FALSE & converted_time != -999 & EVENT_TYPE %in% c("INDOOR", "OUTDOOR", "XC")) %>%
  mutate(
    TIME = converted_time
  ) %>%
  ungroup() %>%
  group_by(EVENT, EVENT_TYPE, GENDER, year_index, week_index) %>%
  summarise(
    week_best_time = min(TIME, na.rm = T)
  ) %>%
  ungroup() %>%
  group_by(GENDER, EVENT, EVENT_TYPE) %>%
  arrange(year_index, week_index, .by_group = TRUE) %>%
  mutate(
    c_week_best_time = cummin(week_best_time)
  ) %>%
  filter(!is.na(year_index)) %>%
  mutate(
    week_key = paste0(year_index, "-", week_index)
  ) %>%
  select(EVENT, EVENT_TYPE, GENDER, week_key, week_best_time, c_week_best_time)

raceBestMonthly <- rl %>%
  filter(is_duplicate == "N") %>%
  filter(IS_FIELD == FALSE & converted_time != -999 & EVENT_TYPE %in% c("INDOOR", "OUTDOOR", "XC")) %>%
  mutate(
    TIME = converted_time
  ) %>%
  ungroup() %>%
  group_by(EVENT, EVENT_TYPE, GENDER, year_index, month_index) %>%
  summarise(
    month_best_time = min(TIME, na.rm = T)
  ) %>%
  ungroup() %>%
  group_by(GENDER, EVENT, EVENT_TYPE) %>%
  arrange(year_index, month_index, .by_group = TRUE) %>%
  mutate(
    c_month_best_time = cummin(month_best_time)
  ) %>%
  mutate(
    month_key = paste0(year_index, "-", month_index)
  ) %>%
  select(EVENT, EVENT_TYPE, GENDER, month_key, month_best_time, c_month_best_time)
```

### Calculate SOC
```{r warning=FALSE}
# Create data for race strength (could get messy with heats...) - SOC = Strength of Competition
raceSOC <- raceGrp %>%
  ungroup() %>%
  group_by(
    MEET_NAME, meet_date, competition_year, GENDER, EVENT, EVENT_TYPE
  ) %>%
  summarise(
    avg_top_time = mean(TIME[meet_evnt_rank <= 3], na.rm = T),
    n_runners = n_distinct(NAME),
    winning_time = min(TIME, na.rm = T)
  ) %>%
  ungroup()

# Join SOC data
soc <- raceSOC %>%
  left_join(cal, by = c("meet_date" = "cal_d")) %>%
  left_join(raceBestWeekly, by = c("week_key", "EVENT_TYPE", "GENDER", "EVENT")) %>%
  left_join(raceBestMonthly, by = c("month_key", "EVENT_TYPE", "GENDER", "EVENT"))

# Create SOC value
soc <- soc %>%
  mutate(
    soc = round(((0.1 * (month_best_time / winning_time)) + ((c_month_best_time / avg_top_time) * 0.2) + ((c_month_best_time / winning_time) * 0.7)), 5) * 100
  ) %>%
  filter(!grepl("XC", EVENT) & competition_year >= 2017) %>%
  filter(!(EVENT_TYPE == "INDOOR" & EVENT == "10K")) %>%
    filter(
    !(EVENT == "Mile" & EVENT_TYPE != "INDOOR") &
    !(EVENT == "3000m" & EVENT_TYPE != "INDOOR") &
    !(EVENT == "3000S" & EVENT_TYPE == "INDOOR") &
    !(EVENT == "1500m" & EVENT_TYPE == "INDOOR") &
    !(MEET_NAME == "WSC Relays and Multi-Events Qualifier" & EVENT == "800m") &
    !(tolower(MEET_NAME) == "michael johnson invitational" & lubridate::ymd(meet_date) == "2021-03-19")
  )
```

```{r warning=FALSE}
# Join SOC data to line level
socJoin <- soc %>%
  ungroup() %>%
  select(MEET_NAME, meet_date, GENDER, EVENT, competition_year, soc, week_best_time, c_week_best_time, month_best_time, c_month_best_time)

raceDets <- raceGrp %>%
  left_join(socJoin, by = c("meet_date", "MEET_NAME", "GENDER", "competition_year", "EVENT")) %>% 
  filter(!is.na(soc)) %>%
  group_by(MEET_NAME, meet_date, GENDER, EVENT) %>%
  mutate(
    weighted_finish = ((max(place_behind_winner) - place_behind_winner) / max(place_behind_winner)),
    weighted_finish_time = (time_behind_winner / min(TIME)),
    weighted_time = 1- ((TIME - month_best_time) / month_best_time),
    weighted_overall_time = 1 - ((TIME - c_month_best_time) / c_month_best_time)
  ) %>%
  mutate(
    scaled_weighted_finish = scale(weighted_finish),
    scaled_weighted_finish_time = scale(weighted_finish_time),
    scaled_weighted_time = scale(weighted_time),
    scaled_weighted_overall_time = scale(weighted_overall_time)
  ) %>%
  ungroup() %>%
  mutate(
    rpi = ((soc / 100) * weighted_finish) * 1 + (1 - weighted_finish_time) * (soc / 100) * 1 + (weighted_overall_time) * 8,
    rpi_v2 = ((soc / 100) * scaled_weighted_finish) * 1 + (1 - scaled_weighted_finish_time) * (soc / 100) * 1 + (scaled_weighted_overall_time) * 8 
  ) %>%
  filter(
    !(EVENT == "Mile" & EVENT_TYPE != "INDOOR") &
    !(EVENT == "10K" & TIME < 1000) &
    !(EVENT == "3000m" & EVENT_TYPE != "INDOOR")
  ) %>%
  mutate(
    race_key = paste0(NAME, "-", EVENT, "-", meet_date, "-", TIME, "-", PRELIM)
  ) %>%
  distinct(race_key, .keep_all = TRUE)
```

## Pivot Out Personal Best Data

```{r warning=FALSE}
# Create all-time grouped data
grpAllTime <- rl %>%
  filter.(IS_FIELD == FALSE) %>%
  summarize.(
    pr = min(converted_time, na.rm = T),
    prs_run = sum(is_pr, na.rm = T),
    avg_time = round(mean(converted_time, na.rm = T), 2),
    times_run = n_distinct(MEET_NAME),
    avg_place = round(mean(PLACE, na.rm = T), 2),
    wins = n_distinct(MEET_NAME[PLACE == 1]),
    .by = c(NAME, TEAM, GENDER, EVENT)
  ) %>%
  mutate.(
    win_pct = round((wins / times_run) * 100, 3)
  )

# Rowwise operations
grpAllTime$clean_pr = sapply(grpAllTime$pr, reformatTimes)
grpAllTime$clean_avg_time = sapply(grpAllTime$avg_time, reformatTimes)

# Convert back to dataframe
grpAllTime <- as.data.frame(grpAllTime)

# Create PR df via pivot approach
prDf <- grpAllTime %>%
  pivot_wider(
    id_cols = c(NAME, GENDER, TEAM),
    names_from = EVENT,
    values_from = c(avg_time, pr, times_run)
  ) 

# Pivot out current year
prDf22 <- grpYrly %>%
  filter(competition_year == "2022") %>%
  pivot_wider(
    id_cols = c(NAME, GENDER, TEAM),
    names_from = EVENT,
    values_from = c(avg_time, pr, times_run)
  ) 
  # right_join(names, by = c("NAME", "TEAM", "GENDER"))
```

# Join Divisions to All Time Grouping

```{r}
# Join
prDf <- prDf %>%
  left_join(divNames, by = c("NAME", "GENDER", "TEAM"))

prDf22 <- prDf22 %>%
  left_join(divNames, by = c("NAME", "GENDER", "TEAM"))
```

# Model Attempt

## Subset D1 Women

```{r}
# All-time
wDf <- prDf %>%
  filter(GENDER == "F") %>%
  select(
    NAME, DIVISION, avg_time_3000m, avg_time_5000m, avg_time_1500m, avg_time_Mile,
    pr_800m, pr_1500m, pr_Mile, pr_3000m, pr_5000m, pr_3000S
  )
```

# Visuals

## 3k vs 5k

```{r}
wDf %>%
  filter(pr_3000m <= 650 & pr_3000S <= 660) %>%
  filter(DIVISION %in% c("D1", "D2", "D3", "NAIA")) %>%
  ggplot(aes(pr_3000m, pr_3000S, colour = DIVISION)) +
  geom_point(size = 5) +
  # geom_smooth(method='lm', formula= y~x) +
  scale_x_continuous(name = "Personal Best (3k)",
                     breaks = c(530, 540, 550, 560, 570, 580, 590, 600, 610, 620, 630, 640),
                     labels = c("8:50", "", "9:10", "", "9:30", "", "9:50", "", "10:10", "", "10:30", "")) +
  scale_y_continuous(name = "Personal Best (3k Steeple)",
                   breaks = c(550, 560, 570, 580, 590, 600, 610, 620, 630, 640, 650),
                   labels = c("9:10", "", "9:30", "", "9:50", "", "10:10", "", "10:30", "", "10:50")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 5, size = 5) +
  ggplot2::labs(
    x = "PR 3k",
    y = "PR Steeple",
    caption = "Data shown for indoor & outdoor track seasons for recent NCAA seasons.",
    title = "NCAA Women 3k vs Steeplchase Performances",
    subtitle = "Minimum 10:50 in 3k and 11:00 in steeple."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    legend.position = "none",
    axis.text.x = element_text(face = "italic"),
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 50, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold"),
  ) +
  facet_wrap(DIVISION ~ .)
```

## 2022

### 3k vs 5k in 2022

```{r}
prDf22 %>%
  filter(GENDER == "F") %>%
  filter(pr_3000m <= 600) %>%
  ggplot(aes(pr_3000m, pr_5000m, colour = DIVISION)) +
  geom_point(size = 5) +
  # geom_smooth(method='lm', formula= y~x) +
  scale_x_continuous(name = "Personal Best (3k)",
                     breaks = c(530, 540, 550, 560, 570, 580, 590, 600),
                     labels = c("8:50", "9:00", "9:10", "9:20", "9:30", "9:40", "9:50", "10:00")) +
  scale_y_continuous(name = "Personal Best (5k)",
                     breaks = c(900, 945, 990, 1035, 1080, 1125, 1170),
                     labels = c("15:00", "15:45", "16:30", "17:15", "18:00", "18:45", "19:30")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 5, size = 5) +
  ggplot2::labs(
    x = "PR 3k",
    y = "PR 5k",
    caption = "Data shown for the 2021-2022 indoor & outdoor track seasons.",
    title = "NCAA Women 3k vs 5k Performances in 2022",
    subtitle = "Showing women with 3K PRs of 10:00 or better."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 50, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold"),
  ) +
  facet_wrap(DIVISION ~ .)
```

### 3k vs 5k Averages

```{r}
prDf22 %>%
  filter(GENDER == "F") %>%
  filter(DIVISION %in% c("D1", "D2", "D3", "NAIA")) %>%
  filter(pr_3000m <= 600) %>%
  ggplot(aes(avg_time_3000m, avg_time_5000m, colour = DIVISION)) +
  geom_point(size = 5) +
  # geom_smooth(method='lm', formula= y~x) +
  scale_x_continuous(name = "Average Mark (3k)",
                     breaks = c(530, 540, 550, 560, 570, 580, 590, 600, 610, 620, 630, 640, 650),
                     labels = c("8:50", "9:00", "9:10", "9:20", "9:30", "9:40", "9:50", "10:00", "10:10", "10:20", "10:30", "10:40", "10:50")) +
  scale_y_continuous(name = "Average Mark (5k)",
                     breaks = c(900, 945, 990, 1035, 1080, 1125, 1170),
                     labels = c("15:00", "15:45", "16:30", "17:15", "18:00", "18:45", "19:30")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 7, size = 7) +
  ggplot2::labs(
    x = "PR 3k",
    y = "PR 5k",
    caption = "Data shown for the roughly the past 5-6 years of results..",
    title = "NCAA Women 3k vs 5k Performances for Recent Seasons",
    subtitle = "Showing women with 3K PRs of 10:00 or better."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 50, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold"),
  )
```

### 3k vs Mile

```{r}
prDf22 %>%
  filter(GENDER == "F") %>%
  filter(pr_3000m <= 650 & pr_3000S <= 660) %>%
  #filter(DIVISION %in% c("D1", "D2", "D3")) %>%
  ggplot(aes(pr_3000m, pr_3000S, colour = DIVISION)) +
  geom_point(size = 5) +
  # geom_smooth(method='lm', formula= y~x) +
  scale_x_continuous(name = "Personal Best (3k)",
                     breaks = c(530, 540, 550, 560, 570, 580, 590, 600, 610, 620, 630, 640),
                     labels = c("8:50", "", "9:10", "", "9:30", "", "9:50", "", "10:10", "", "10:30", "")) +
  scale_y_continuous(name = "Personal Best (3k Steeple)",
                   breaks = c(550, 560, 570, 580, 590, 600, 610, 620, 630, 640, 650),
                   labels = c("9:10", "", "9:30", "", "9:50", "", "10:10", "", "10:30", "", "10:50")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 5, size = 5) +
  ggplot2::labs(
    x = "PR 3k",
    y = "PR Steeple",
    caption = "Data shown for the 2021-2022 indoor & outdoor track seasons.",
    title = "NCAA Women 3k vs Steeplchase Performances in 2022",
    subtitle = "Minimum 10:50 in 3k and 11:00 in steeple."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    legend.position = "none",
    axis.text.x = element_text(face = "italic"),
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 50, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=30, face = "bold"),
  ) +
  facet_wrap(DIVISION ~ .)
```

## 800 v Mile

```{r}
wDf %>%
  filter(pr_Mile <= 300 & pr_800m <= 160) %>%
  filter(DIVISION != "OTHER") %>%
  ggplot(aes(pr_Mile, pr_800m, colour = DIVISION)) +
  geom_point() +
  scale_y_continuous(name = "Personal Best (800m)",
                     breaks = c(115, 120, 125, 130, 135, 140, 145, 150, 155, 160),
                     labels = c("1:55", "2:00", "2:05", "2:10", "2:15", "2:20", "2:25", "2:30", "2:35", "2:40")) +
  scale_x_continuous(name = "Personal Best (Mile)",
                     breaks = c(255, 265, 275, 285, 295, 305),
                     labels = c("4:15", "4:25", "4:35", "4:45", "4:55", "5:05")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 10) +
  ggplot2::labs(
    x = "PR 800",
    y = "PR Mile",
    caption = "Data broken out by competition division.",
    title = "NCAA Women 800 vs Mile PRs",
    subtitle = "Minimum 5:00 mile & 2:40 800m."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 25, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold"),
  ) +
  facet_wrap(DIVISION ~ .)
```

## 5k vs Mile

```{r}
wDf %>%
  filter(pr_Mile <= 330 & pr_5000m <= 1170) %>%
  ggplot(aes(pr_Mile, pr_5000m, colour = DIVISION)) +
  geom_point(size = 3) +
  # geom_smooth(method='lm', formula= y~x) +
  scale_x_continuous(name = "Personal Best (Mile)",
                     breaks = c(255, 265, 275, 285, 295, 305, 315, 325),
                     labels = c("4:15", "4:25", "4:35", "4:45", "4:55", "5:05", "5:15", "5:25")) +
  scale_y_continuous(name = "Personal Best (5k)",
                     breaks = c(900, 945, 990, 1035, 1080, 1125, 1170),
                     labels = c("15:00", "15:45", "16:30", "17:15", "18:00", "18:45", "19:30")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 10) +
  ggplot2::labs(
    caption = "",
    title = "NCAA Women Mile vs 5k PRs",
    subtitle = "Showing women with 3K PRs of 10:00 or better."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 25, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold"),
  ) +
  facet_wrap(DIVISION ~ .)
```

## 5k Improvement

```{r}
grpLines %>%
  filter(GENDER == "F" & DIVISION == "D1") %>%
  filter(YEAR == 2021 & EVENT == "5000m" & PR <= 990) %>%
  filter(!is.na(PR_PROG)) %>%
  ggplot(aes(PR, PR_PROG)) +
  geom_point() +
  scale_x_continuous(name = "Personal Best", 
                     breaks = c(900, 925, 950, 975), 
                     labels = c("15:00", "15:25", "15:50", "16:15")) +
  geom_label_repel(aes(label = NAME), max.overlaps = 200, colour = "blue") +
  theme_minimal() +
  theme(legend.position = "none", text = element_text(size=20)) +
  ggtitle("D1 Women 5k Progession in 2021", subtitle = "PR improvement in 5k vs previous years. Only showing for PRs < 16:31.") +
  labs(y = "PR Improvement (Sec)")
```

## PR 3k vs 5k

```{r}
prDf %>%
  filter(DIVISION == "D1" & GENDER == "F" & PR_3000m <= 570) %>%
  ggplot(aes(PR_3000m, PR_5000m)) +
  geom_point() +
  stat_smooth() +
  geom_label_repel(aes(label = NAME), max.overlaps = 100)
```

## Overall Grouped Data

```{r}
grpYrly %>%
  filter(GENDER == "F" & competition_year == "2022") %>%
  ggplot(aes(pr, avg_time, colour = DIVISION)) +
  geom_point(size = 3) +
  # geom_smooth(method='lm', formula= y~x) 
  geom_label_repel(aes(label = NAME), max.overlaps = 10) +
  ggplot2::labs(
    x = "PR 3k",
    y = "PR 5k",
    caption = "",
    title = "NCAA Women 3k vs 5k PRs",
    subtitle = "Showing women with 3K PRs of 10:00 or better."
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold"),
    plot.title.position = "panel",
    text = element_text(size=20),
    #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    legend.position = "none"
  ) +
  facet_wrap(DIVISION ~ .)
```

# 6k XC Wins

```{r}
grpLines %>%
  filter(GENDER == "F" & EVENT == "6K XC" & YEAR == 2021 & AVG_TIME >= 1000) %>%
  slice_max(n = 50, order_by = PR) %>%
  ggplot(aes(AVG_TIME, AVG_PLACE, colour = DIVISION)) +
  geom_point() +
  geom_label_repel(aes(label = NAME), max.overlaps = 20)
  ggplot2::labs(
    x = "",
    y = "",
    caption = "",
    title = "Top 50 NCAA Women Based on 6k XC Win %",
    subtitle = "Data shown for top 50 women with most wins."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=10, face = "bold")
  )
  # scale_x_continuous(name = "Win Percentage",
  #                    breaks = c(0.0, 20.0, 40.0, 60.0, 80.0, 100.0),
  #                    labels = c("0%", "20%", "40%", "60%", "80%", "100%"))
```

# Post PR Analysis

## Subset Data

```{r}
prs <- lines %>%
  filter(post_pr_shift != -999 & EVENT_TYPE == "INDOOR" & PRELIM == FALSE) %>%
  filter(EVENT %in% c("800m", "Mile", "3000m", "5000m")) %>%
  filter(
    !(EVENT == "800m" & abs(post_pr_shift) >= 70) &
    !(EVENT == "Mile" & abs(post_pr_shift) >= 120) &
    !(EVENT == "3000m" & abs(post_pr_shift) >= 200) &
    !(EVENT == "5000m" & abs(post_pr_shift) >= 300)
  )
```

## Plot Post-PR Performance

```{r}
prs %>%
  filter(DIVISION %in% c("D1")) %>%
  ggplot(aes(post_pr_shift, fill = GENDER)) +
  geom_density(alpha = 0.5) +
  # geom_label_repel(aes(label = NAME), max.overlaps = 20)
  ggplot2::labs(
    x = "Time Shift Post-PR (seconds)",
    y = "Proportion of Performances",
    caption = "Only using Division 1 runners.",
    title = "Distribution of Post-PR Performances Relative to PR",
    subtitle = "Time shift is the amount (in seconds) a runner was better/worse in an event the next time they competed after setting a PR. 
    Ex: A 4:02 miler running 3:59 would be +3"
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=10, face = "bold")
  ) +
  facet_wrap(EVENT ~ ., scales = "free")
```

## Plot For Big Jumps in Performance

```{r}
prs %>%
  filter(!is.na(TIME_SHIFT)) %>%
  filter(DIVISION %in% c("D1")) %>%
  filter(
    (TIME_SHIFT >= 2 & EVENT == "800m") |
    (TIME_SHIFT >= 5 & EVENT == "Mile") |
    (TIME_SHIFT >= 10 & EVENT == "3000m") |
    (TIME_SHIFT >= 20 & EVENT == "5000m")
  ) %>%
  ggplot(aes(post_pr_shift, fill = GENDER)) +
  geom_density(alpha = 0.5) +
  # geom_label_repel(aes(label = NAME), max.overlaps = 20)
  ggplot2::labs(
    x = "Time Shift Post-PR (seconds)",
    y = "Proportion of Performances",
    caption = "Only using Division 1 runners.",
    title = "Distribution of Post-PR Performances Relative to PR for Big Improvements",
    subtitle = "Time shift is the amount (in seconds) a runner was better/worse in an event the next time they competed after setting a PR. Ex: A 4:02 miler running 3:59 would be +3
    Big improvements defined as 2+ sec in 800, 5+ in mile, 10+ in 3k, and 20+ in 5k."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    # legend.position = "none",
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=10, face = "bold")
  ) +
  facet_wrap(EVENT ~ ., scales = "free")
```

# 2021 Improvements

```{r}
lines %>%
  ungroup() %>%
  filter(!is.na(TIME_SHIFT) & GENDER == "F" & EVENT == "5000m") %>%
  filter(YEAR == 2021 & EVENT_TYPE == "INDOOR" & is_pr == 1) %>%
  filter(MEET_DATE >= "2021-12-01") %>%
    filter(
    !(EVENT == "800m" & abs(TIME_SHIFT) >= 70) &
    !(EVENT == "Mile" & abs(TIME_SHIFT) >= 120) &
    !(EVENT == "3000m" & abs(TIME_SHIFT) >= 200) &
    !(EVENT == "5000m" & abs(TIME_SHIFT) >= 300)
  ) %>%
  filter(TIME <= 1000) %>%
  #filter(DIVISION %in% c("D1", "D2", "D3")) %>%
  dplyr::slice_max(n = 25, order_by = TIME_SHIFT) %>%
  ggplot(aes(reorder(NAME, TIME_SHIFT), TIME_SHIFT, fill = DIVISION)) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = clean_time), fill = "white") +
  ggplot2::labs(
    x = "",
    y = "Improvement on Previous Personal Best (Seconds)",
    caption = "Some transfers may be omitted due to data issues.",
    title = "Top 25 Improvements for 21-22 Indoor Season vs Personal Best - Men's 5k",
    subtitle = "Runners without a mark are not included. Label shows runners new personal best mark."
  ) +
  ggplot2::theme_bw() +
  scale_fill_stata() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold")
  ) + 
  coord_flip()
```

# Top 2021 Performances

## Avg Time

```{r}
runnerGrpYrly %>%
  ungroup() %>%
  filter(!is.na(AVG_TIME) & GENDER == "F" & EVENT == "5000m") %>%
  filter(YEAR == 2021) %>%
  #filter(DIVISION %in% c("D1", "D2", "D3")) %>%
  dplyr::slice_min(n = 25, order_by = AVG_TIME) %>%
  ggplot(aes(reorder(NAME, -AVG_TIME), AVG_TIME, fill = DIVISION)) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = clean_avg_time), fill = "white") +
  ggplot2::labs(
    x = "",
    y = "Improvement on Previous Personal Best (Seconds)",
    caption = "Some transfers may be omitted due to data issues.",
    title = "Top 25 Improvements for 21-22 Indoor Season vs Personal Best - Men's 5k",
    subtitle = "Runners without a mark are not included. Label shows runners new personal best mark."
  ) +
  ggplot2::theme_bw() +
  scale_fill_stata() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold")
  ) + 
  coord_flip()
```

# Top Runner Progression

## Get Top Runners

```{r}
# Set event and gender
evnt <- "Mile"
gender <- "F"
type <- "INDOOR"

topNames <- grpYrly %>%
  ungroup() %>%
  filter(competition_year %in% c(2021, 2022)) %>%
  filter(GENDER == gender & EVENT == evnt & EVENT_TYPE == type) %>%
  slice_min(n = 15, order_by = pr)

top <- grpYrly %>%
  filter(competition_year >= 2017) %>%
  filter(EVENT == evnt & EVENT_TYPE == type & GENDER == gender) %>%
  filter(NAME %in% topNames$NAME)

top %>%
  ggplot(aes(competition_year, avg_place, color = NAME, group = NAME)) +
  geom_point() +
  geom_line() +
  geom_label_repel(data = top[which(top$competition_year == 2019), ], aes(label = NAME)) +
  # geom_label(aes(label = clean_avg_time), fill = "white") +
  ggplot2::labs(
    x = "",
    y = "Improvement on Previous Personal Best (Seconds)",
    caption = "Some transfers may be omitted due to data issues.",
    title = "Top 25 Improvements for 21-22 Indoor Season vs Personal Best - Men's 5k",
    subtitle = "Runners without a mark are not included. Label shows runners new personal best mark."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 20, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold"),
    legend.position = "none"
  )
```

# SOC Meets

## Women's Table

```{r}
pal_hex <- c("#762a83", "#af8dc3", "#e7d4e8", "#f7f7f7",
             "#d9f0d3", "#7fbf7b", "#1b7837")

rows <- 35
div <- "D1"

tab <- soc %>%
  ungroup() %>%
  funique() %>%
  filter(GENDER == "F" & competition_year > 2015) %>%
  filter(MEET_NAME != "west coast relays") %>%
  filter(EVENT %in% c("800m", "Mile", "1500m", "3000m", "5000m", "10k", "3000S")) %>%
  slice_max(n = rows, order_by = soc) %>%
  select(
    MEET_NAME, meet_date, EVENT, soc
  ) %>%
  gt::gt() %>% 
  cols_label(MEET_NAME = "Meet", meet_date = "Date", EVENT = "Event", soc = "Strength of Competition (SOC)") %>% 
  gt::data_color(
    columns = soc, 
    colors = scales::col_numeric(
      palette = pal_hex,
      domain = range(soc)
      )
    ) %>% #, colors = hulk_pal) %>%
  gtExtras::gt_theme_538() %>%
  gt::tab_header(title = paste(glue::glue("Top {rows} Competitions by NCAA Women Since 2019 Based on SOC"))) %>%
  gt::tab_source_note(gt::md(glue::glue('**Notes**: @ivanxcky | SOC is Strength of Competition | Data as of {lubridate::today()}'))) %>%
  tab_options(data_row.padding = px(2)) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = c(1, 3, 4))
  ) %>%
  fmt_number(columns = c(4), decimals = 3) %>%
  fmt_missing(columns = everything(), rows = everything(), missing_text = " ") %>%
  tab_style(
    style = list(
      cell_fill(color = "white")
    ),
    locations = cells_body(
      columns = 4,
      rows = ((rows+1)/2)
    )
  )

tab

tab %>%
  gtsave("w_soc_all.png")
```

## Men's Table

```{r}
tab <- soc %>%
  ungroup() %>%
  funique() %>%
  filter(GENDER == "M" & competition_year > 2018) %>%
  filter(MEET_NAME != "west coast relays") %>%
  filter(EVENT %in% c("800m", "Mile", "1500m", "3000m", "5000m", "10k", "3000S")) %>%
  slice_max(n = rows, order_by = soc) %>%
  select(
    MEET_NAME, meet_date, EVENT, soc
  ) %>%
  gt::gt() %>% 
  cols_label(MEET_NAME = "Meet", meet_date = "Date", EVENT = "Event", soc = "Strength of Competition (SOC)") %>% 
  gt::data_color(
    columns = soc, 
    colors = scales::col_numeric(
      palette = pal_hex,
      domain = range(soc)
      )
    ) %>% #, colors = hulk_pal) %>%
  gtExtras::gt_theme_538() %>%
  gt::tab_header(title = paste(glue::glue("Top {rows} Competitions by NCAA Men Since 2019 Based on SOC"))) %>%
  gt::tab_source_note(gt::md(glue::glue('**Notes**: @ivanxcky | SOC is Strength of Competition | Data as of {lubridate::today()}'))) %>%
  tab_options(data_row.padding = px(2)) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = c(1, 3, 4))
  ) %>%
  fmt_number(columns = c(4), decimals = 3) %>%
  fmt_missing(columns = everything(), rows = everything(), missing_text = " ") %>%
  tab_style(
    style = list(
      cell_fill(color = "white")
    ),
    locations = cells_body(
      columns = 4,
      rows = ((rows+1)/2)
    )
  )

tab

tab %>%
  gtsave("m_soc_all.png")
```

# RPI

```{r}
raceDets %>%
  mutate(
    rpi = round(rpi, 3)
  ) %>%
  filter(
   DIVISION == "D1" & GENDER == "F" & competition_year >= 2018 & EVENT %in% c("Mile", "800m", "3000m", "5000m", "1500m", "10k", "3000S") & 
     EVENT_TYPE == "OUTDOOR"
  ) %>%
  ungroup() %>%
  slice_max(n = 15, order_by = rpi) %>%
  ggplot(aes(reorder(paste0(NAME, " - ", TEAM, " - ", MEET_NAME, " (", EVENT, ")"), rpi), rpi, fill = as.factor(EVENT))) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = rpi), fill = "white", size = 8) +
  ggplot2::labs(
    x = "",
    y = "RPI",
    caption = "Data only for 2021-2022 indoor racing season.",
    title = "Top 15 NAIA Women's Races of 2022 Indoor Season",
    subtitle = "Races are based on RPI (Race Performance Index)."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 45, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=35, face = "bold")
  ) +
  guides(fill=guide_legend(title="Event")) +
  coord_flip()
```

# Table of RPI

```{r}
pal_hex <- c("#762a83", "#af8dc3", "#e7d4e8", "#f7f7f7",
             "#d9f0d3", "#7fbf7b", "#1b7837")

rows <- 50
div <- "D1"

tab <- raceDets %>%
  mutate(
    rpi = round(rpi, 3)
  ) %>%
  filter(
   DIVISION == div & GENDER == "F" & competition_year >= 2016 & EVENT %in% c("Mile", "800m", "3000m", "5000m", "1500m", "10k", "3000S")
  ) %>%
  ungroup() %>%
  slice_max(n = rows, order_by = rpi) %>%
  select(
    NAME, TEAM, DIVISION, MEET_NAME, meet_date, EVENT, PLACE, MARK, rpi
  ) %>%
  gt::gt() %>% 
  cols_label(NAME = "Athlete", TEAM = "School", DIVISION = "Division", MEET_NAME = "Meet", 
             meet_date = "Date", EVENT = "Event", PLACE = "Place", MARK = "Time", rpi = "RPI") %>% 
  gt::data_color(
    columns = rpi, 
    colors = scales::col_numeric(
      palette = pal_hex,
      domain = range(rpi)
      )
    ) %>% #, colors = hulk_pal) %>%
  gtExtras::gt_theme_538() %>%
  gt::tab_header(title = paste(glue::glue("Top {rows} Performances by NCAA {div} Women Since 2016 Based on RPI"))) %>%
  gt::tab_source_note(gt::md(glue::glue('**Notes**: @ivanxcky | RPI is Race Performance Index | Data as of {lubridate::today()}'))) %>%
  tab_options(data_row.padding = px(2)) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = c(1:2, 6, 8:9))
  ) %>%
  fmt_number(columns = c(9), decimals = 3) %>%
  fmt_missing(columns = everything(), rows = everything(), missing_text = " ") %>%
  tab_style(
    style = list(
      cell_fill(color = "white")
    ),
    locations = cells_body(
      columns = 9,
      rows = ((rows+1) / 2) - 0.5
    )
  )

tab

tab %>%
  gtsave("w_D1_rpi_16_22_all.png")
```

# Best Runners of 2021

```{r}
raceDets %>%
  mutate(
    wsor = round(wsor, 4)
  ) %>%
  filter(
   DIVISION == "NAIA" & GENDER == "F" & competition_year == 2022 & EVENT_TYPE == "INDOOR" & EVENT %in% c("Mile", "800m", "3000m", "5000m")
  ) %>%
  group_by(NAME, TEAM) %>%
  summarise(
    n_races = n_distinct(race_key),
    avg_wsor = round(mean(wsor, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  filter(n_races >= 2) %>%
  slice_max(n = 30, order_by = avg_wsor) %>%
  ggplot(aes(reorder(NAME, avg_wsor), avg_wsor, fill = as.factor(TEAM))) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = avg_wsor), fill = "white", size = 3.5) +
  ggplot2::labs(
    x = "",
    y = "Avg. WSOR",
    caption = "Data through 2.12.22",
    title = "Top 30 NAIA Women of 2022 Indoor Season",
    subtitle = "Races are based on WSOR (Weighted Score of Run)."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    # axis.text.x = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 25, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=18, face = "bold"),
    legend.position = "none"
  ) +
  guides(fill=guide_legend(title="Event")) +
  coord_flip()
```

# Best Performances of 2021

```{r}
raceDets %>%
  mutate(
    wsor = round(wsor, 4)
  ) %>%
  filter(
   DIVISION == "D2" & GENDER == "F" & competition_year == 2022 & EVENT_TYPE == "INDOOR" & EVENT %in% c("Mile", "800m", "3000m", "5000m")
  ) %>%
  group_by(NAME, TEAM) %>%
  summarise(
    n_races = n_distinct(race_key),
    best_wsor = round(max(wsor, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  filter(n_races >= 1) %>%
  slice_max(n = 30, order_by = best_wsor) %>%
  ggplot(aes(reorder(NAME, best_wsor), best_wsor, fill = as.factor(TEAM))) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = best_wsor), fill = "white", size = 3.5) +
  ggplot2::labs(
    x = "",
    y = "WSOR",
    caption = "Data through 2.4.22",
    title = "Top 30 D2 Women of 2022 Indoor Season",
    subtitle = "Races are based on WSOR (Weighted Score of Run)."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    # axis.text.x = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 25, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=18, face = "bold"),
    legend.position = "none"
  ) +
  guides(fill=guide_legend(title="Event")) +
  coord_flip()
```

```{r}
raceDets %>%
  mutate(
    wsor = round(wsor, 4)
  ) %>%
  filter(
   DIVISION == "D1" & GENDER == "F" & competition_year >= 2019 & EVENT == "5000m"
  ) %>%
  ungroup() %>%
  slice_min(n = 25, order_by = converted_time) %>%
  ggplot(aes(reorder(paste0(NAME, " - ", TEAM, " - ", MEET_NAME, " (", competition_year, ")"), wsor), wsor, fill = as.factor(competition_year))) +
  geom_bar(stat = "identity") +
  geom_label(aes(label = MARK), fill = "white", size = 2) +
  ggplot2::labs(
    x = "",
    y = "WSOR",
    caption = "Omitting XC competitions.",
    title = "Top 25 D1 Women's Races of 2021 Season",
    subtitle = "Races are based on WSOR (Weighted Score of Run)."
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    aspect.ratio = 9 / 16,
    axis.text.y = element_text(face = "italic"),
    plot.title = ggplot2::element_text(size = 22, hjust = 0.5, vjust = 0.5, face = "bold"),
    plot.subtitle = element_text(face = "italic"),
    text = element_text(size=12, face = "bold")
  ) +
  guides(fill=guide_legend(title="Event")) +
  coord_flip()
```
